Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2018-08-07 paper/manuscript.tex
updated manuscript.

\section{Step 3 -- Overview of available analysis techniques}
\label{sec:step3}
\todo[inline, color={red!40}]{JDC, BA: uncertainty estimation section}
\begin{enumerate}
\item Detecting boundary between equilibrated and production regions (Chodera: \url{http://dx.doi.org/10.1021/acs.jctc.5b00784})
\item Decorrelating samples for analysis
\begin{itemize}
\item Subsample different lambdas based on correlation times
\item Ensure all simulations at least 50x correlation time %MRS: I've heard 40x from Swope. Rationale has to do with the number of uncorrelated samples for the estimate of the uncertainty to be reliable.
\end{itemize}
\item Examining output data for common problems with discussions of what exactly to plot or look at; examples of typical curves for dV/dlambda and free energy versus lambda, for example
\begin{itemize}
\item Make sure ligand doesn’t tumble out of binding site (Mey has observed this)
\item If possible, look at multiple estimators.  Significant discrepancies between different free energy estimators (TI, BAR, MBAR) indicate there is a problem..
\item Poor replica mixing (for replica-exchange)
\item Correlation time as a function of lambda as it would be expected to be a smooth
\item Dependence on initial conformation (best practice would be to look at 
\item Torsional analysis: Is it stuck in specific states? Only very rarely transitions?
\item More ``usual suspects''
\end{itemize}
\item Estimators for free energies
\begin{itemize}
\item MBAR is gold standard, but requires energies from each sample at all of the other states.  BAR is almost as good if the states are separated so as to minimize the number of states (i.e. there is good overlap from $i \rightarrow i+1$ and $i \rightarrow i-1$, but not to $i\pm 2$.
\item Using bootstrap (or block bootstrap) for errors is more reliable than analytical error estimates. BAR uncertainties are underestimate.
\item TI should be roughly in agreement, but quadrature error hard to quantify; should generally be used only if the path is known to be good for that size of molecule.
\item Other variants useful in special circumstances (e.g. Z. Tan stochastic version)
\item expontential averaging is very problematic.  If using intermediates (which will almost always be the case) should avoid.
\end{itemize}
\item Computing and reporting uncertainties on free energies
\item correlated bootstrap versus timeseries subsampling.
\begin{itemize}
\item Quantifying standard error in dG estimate
\item Variance in dG estimate using multiple methods (TI, DEXP, IEXP, BAR, MBAR, GDEL, etc.)
\item Agreement in dG estimate when repeating calculations with different parameters (random seeds, initial configurations, forcefield, etc.)
\item looking at the overlap matrix and making sure that there is decent overlap between all states.
\item Calculating differences in free energy change as a function of time
\begin{itemize}
	\item Starting from beginning or end of simulation
	\item Significance of differences in midpoint estimate (High middle error: high uncertainty)
\end{itemize}
\item Ensemble method to combine uncertainties into interpretable weighted metric
\begin{itemize}
	\item Simple normalized metric to determine confidence in calculation
	\item Easily interpreted by chemist/biologist when prioritizing new chemistries
\end{itemize}
\end{itemize}
\item Other considerations for many transformations
\begin{itemize}
Compute cycle closure error
\end{itemize}

>>>>

\section{Step 3 -- Overview of available analysis techniques}
\label{sec:step3}
\todo[inline, color={red!40}]{JDC, BA: uncertainty estimation section}
\begin{enumerate}
\item Detecting boundary between equilibrated and production regions (Chodera: \url{http://dx.doi.org/10.1021/acs.jctc.5b00784})
\item Decorrelating samples for analysis
\begin{itemize}
\item Subsample different lambdas based on correlation times
\item Ensure all simulations at least 50x correlation time %MRS: I've heard 40x from Swope. Rationale has to do with the number of uncorrelated samples for the estimate of the uncertainty to be reliable.
\end{itemize}
\item Examining output data for common problems with discussions of what exactly to plot or look at; examples of typical curves for dV/dlambda and free energy versus lambda, for example
\begin{itemize}
\item Make sure ligand doesn’t tumble out of binding site (Mey has observed this)
\item If possible, look at multiple estimators.  Significant discrepancies between different free energy estimators (TI, BAR, MBAR) indicate there is a problem..
\item Poor replica mixing (for replica-exchange)
\item Correlation time as a function of lambda as it would be expected to be a smooth
\item Dependence on initial conformation (best practice would be to look at 
\item Torsional analysis: Is it stuck in specific states? Only very rarely transitions?
\item More ``usual suspects''
\end{itemize}
\item Estimators for free energies
\begin{itemize}
\item MBAR is gold standard, but requires energies from each sample at all of the other states.  BAR is almost as good if the states are separated so as to minimize the number of states (i.e. there is good overlap from $i \rightarrow i+1$ and $i \rightarrow i-1$, but not to $i\pm 2$.
\item Using bootstrap (or block bootstrap) for errors is more reliable than analytical error estimates. BAR uncertainties are underestimate.
\item TI should be roughly in agreement, but quadrature error hard to quantify; should generally be used only if the path is known to be good for that size of molecule.
\item Other variants useful in special circumstances (e.g. Z. Tan stochastic version)
\item expontential averaging is very problematic.  If using intermediates (which will almost always be the case) should avoid.
\item Looking at the overlap matrix and making sure that there is decent overlap between all states to ensure that there are sufficiently high number of intermediates.
\end{itemize}
\item Computing and reporting uncertainties on free energies
\begin{itemize}
\item Using correlated bootstrap versus timeseries subsampling.
\item Variation in dG estimate using multiple methods (TI, DEXP, IEXP, BAR, MBAR, GDEL, etc.)
\item Agreement in dG estimate when repeating calculations with different parameters (random seeds, initial configurations, forcefield, etc.)
\item Calculating differences in free energy change as a function of time
\begin{itemize}
	\item Starting from beginning or end of simulation
	\item Significance of differences in midpoint estimate (High middle error: high uncertainty)
\end{itemize}
\item Ensemble method to combine uncertainties into interpretable weighted metric
\begin{itemize}
	\item Simple normalized metric to determine confidence in calculation
	\item Easily interpreted by chemist/biologist when prioritizing new chemistries
\end{itemize}
\end{itemize}

\item Other considerations for many transformations
\begin{itemize}
\item Compute cycle closure error
\end{itemize}
